A generic and fast C++ optimization framework

### 2017-10-21 paper.tex
Add simple optimizer class.

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{tabularx}
\usepackage{threeparttable}
\usepackage{minted}
\usepackage{caption}
\usepackage{parcolumns}

@@

This section details the usage of mlpack's optimization framework. In this
example, we would like to minimize a simple sparse function, where each
dimension has a parabola with a distinct minimum. In this example, we will show
how we can use mlpack’s framework for this task, and obtain the objective of the
sparse function shown below.

\begin{figure}[H]
\begin{minipage}[t]{0.49\textwidth}
\centering
\begin{minted}[fontsize=\footnotesize]{c++}
class ObjectiveFunction {
 ObjectiveFunction() {
  in = arma::vec(
    "20 12 15 100");
  bi = arma::vec("-4 -2 -3 -8");
 }

 void Shuffel() {
  order = arma::shuffle(
   arma::linspace<arma::uvec>(0, 3, 4));
 }

@@

 double Evaluate(const arma::mat& p,
  size_t s, size_t b) {
  double cost = 0;
  for (size_t i = s; i < s + b; i++) {
   cost += std::pow(p(i), 2) +
    bi(order(i)) * p(order(i)) +
    in(order(i));
  }
  return cost;
 }

 void Gradient(const arma::mat& p,
  size_t s, size_t b, arma::sp_mat& g) {
  g.zeros(arma::size(p));

  for (size_t i = s; i < s + b; i++) {
   g(i) += 2 * p(i) + bi(order(i));
 }
}
\end{minted}
\vspace{2\baselineskip}
% \captionof{listing}{Distinct parabola sparse function.}
\end{minipage}
\begin{minipage}[t]{0.49\textwidth}
\centering
\begin{minted}[fontsize=\footnotesize]{c++}
class ObjectiveFunction {
 ObjectiveFunction() {
  in = arma::vec(
    "20 12 15 100");
  bi = arma::vec("-4 -2 -3 -8");
 }

 void Shuffel() {
  order = arma::shuffle(
   arma::linspace<arma::uvec>(0, 3, 4));
 }

 double Evaluate(const arma::mat& p,
  size_t s, size_t b) {
  double cost = 0;
  for (size_t i = s; i < s + b; i++) {
   cost += std::pow(p(i), 2) +
    bi(order(i)) * p(order(i)) +
    in(order(i));
  }
  return cost;
 }

 void Gradient(const arma::mat& p,
  size_t s, size_t b, arma::sp_mat& g) {
  g.zeros(arma::size(p));

  for (size_t i = s; i < s + b; i++) {
   g(i) += 2 * p(i) + bi(order(i));
 }
}
\end{minted}
% \captionof{listing}{Optimizer class.}
\end{minipage}
\captionof{listing}{not relevant}
\end{figure}

>>>>

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{tabularx}
\usepackage{threeparttable}
\usepackage[newfloat]{minted}
\usepackage{caption}
\usepackage{parcolumns}

@@

This section details the usage of mlpack's optimization framework. In this
example, we would like to minimize a simple function, where each dimension has a
parabola with a distinct minimum. In this example, we will show how we can use
mlpack’s framework for this task, and obtain the objective of the function shown
below.

\begin{listing}[H]
\begin{minted}[fontsize=\footnotesize]{c++}
class ObjectiveFunction {
 ObjectiveFunction() {
  in = arma::vec("20 12 15 100");
  bi = arma::vec("-4 -2 -3 -8");
 }

 void Shuffel() { order = arma::shuffle(arma::linspace<arma::uvec>(0, 3, 4)); }

@@

 double Evaluate(const arma::mat& para, size_t s, size_t bs) {
  double cost = 0;
  for (size_t i = s; i < s + bs; i++)
   cost += std::pow(para(i), 2) + bi(order(i)) * para(order(i)) + in(order(i));
  return cost;
 }

 void Gradient(const arma::mat& para, size_t s, size_t bs, arma::mat& g) {
  g.zeros(arma::size(p));
  for (size_t i = s; i < s + bs; i++)
   g(i) += 2 * para(i) + bi(order(i));
  g /= bs;
 }

 arma::vec in, bi; arma::uvec order;
}
\end{minted}
\caption{Simple objective function where each dimension has a parabola with a
  distinct minimum}
\end{listing}
\begin{listing}[H]
\begin{minted}[fontsize=\footnotesize]{c++}
class SimpleOptimizer {
 SimpleOptimizer(size_t bs = 2, double lr = 0.1) : bs(bs), lr(lr) {}

 template<typename FunctionType>
 double Optimize(FunctionType& function, arma::mat& parameter) {
  arma::mat gradient;
  for (size_t i = 0; i < 300; ++i)
  {
   const size_t funIdx = i % 4;
   function.Gradient(parameter, funIdx, gradient, batchSize);
   parameter -= lr * gradient;

   function.Shuffel();
  }

  return function.Evaluate(parameter, 0, function.NumFunctions());
 }
 size_t bs;  double lr;
}
\end{minted}
\caption{Simple stochastic gradient descent method with a static number of
iterations to optimize the function parameter ({\tt bs} defines the batch size
and {\tt lr} the learning rate). )}
\end{listing}
